---
title: "Characterizing Topics with Background Noise"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Characterizing Topics with Background Noise}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  K: 5
  N: 30
  V: 50
  alpha: 0
  id: 1
  method: "product"
  n_models: 5
  out_dir: "gradient"
  save: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(params$id)
```

This vignette demonstrates the use of alignment to characterize topics that are
contaminated by a kind of background noise. It corresponds to the second
simulation in the manuscript accompanying this package. The arguments to this
vignette (which can be modified in the original rmarkdown's YAML) are,

* `K`: The true number of topics underlying the simulated data. In the
manuscript, $K = 5$.
* `N`: The number of samples (i.e., documents) to simulate. In the manuscript, this is set to 250.
* `V`: The number of dimensions (i.e. vocabulary size) per sample. In the manuscript, this is set to 1000.
* `alpha`: To what extent are the simulated data from a true LDA model (as
opposed to background noise). Set to 1 for data from an LDA model and 0 for pure
background noise.
* `id`: A descriptive ID to associate with any saved results.
* `method`: The alignment strategy to pass to `align_topics`.
* `n_models`: The total number of models to fit to the simulated data. In the
manuscript, this is set to 10.
* `out_dir`: If results are saved, where should they be saved to?
* `save`: Should any results be saved?

We load packages. The `sim_gradient` function that generates the contaminated
LDA model is sourced from the link below.

```{r}
library(alto)
library(dplyr)
library(purrr)
library(readr)
library(stringr)
source("https://raw.githubusercontent.com/krisrs1128/topic_align/main/simulations/simulation_functions.R")
my_theme()
```

The block below simulates from the contaminated LDA model. The $\lambda_{\beta}$
and $\lambda_{\gamma}$ parameters are as in the LDA simulation in the
`sim_lda.Rmd` vignette. The $\lambda_{\nu}$ parameter specifies the
hyperparameter of the Dirichlet distribution used in the background noise.

```{r}
attach(params)
lambdas <- list(beta = 0.1, gamma = 0.5, nu = 0.5, count = 1e4)
sim_data <- simulate_gradient(N, K, V, lambdas, alpha = params$alpha)
```

Next, we run a sequence of LDA models and compute an alignment. This code is
identical across vignettes.

```{r}
lda_params <- map(1:n_models, ~ list(k = .))
names(lda_params) <- str_c("K", 1:n_models)
alignment <- sim_data$x %>%
  run_lda_models(lda_params, reset = TRUE) %>%
  align_topics(method = params$method)
```

Next, we extract the data that summarize the quality of topics emerging from the
alignment.

```{r}
scores <- topics(alignment)  %>%
  mutate(id = params$id)
key_topics <- compute_number_of_paths(alignment)
```

By saving these data into different directories, we can gather results across a
variety of $\alpha$'s.

```{r}
id_vars <- params[c("out_dir", "method", "alpha", "id", "N", "V", "K")]
if (params$save) {
  dir.create(params$out_dir, recursive = TRUE)
  write_csv(scores, save_str("scores", id_vars))
  write_csv(key_topics, save_str("key_topics", id_vars))
  exper <- list(sim_data, alignment)
  save(exper, file = save_str("exper", id_vars, "rda"))
}
```
