---
title: "Identifying True Topics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Identifying True Topics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  K: 3
  N: 30
  V: 50
  id: 1
  method: "product"
  n_models: 5
  out_dir: "lda"
  save: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>") 
set.seed(params$id)
```

This vignette applies topic alignment to data generated from a true LDA model.
This corresponds to the first simulation in the manuscript accompanying this
package. The arguments to this vignette (which can be modified in the original
rmarkdown's YAML) are,

* `K`: The true number of topics underlying the simulated data. In the
manuscript, $K = 5$.
* `N`: The number of samples (i.e., documents) to simulate. In the manuscript, this is set to 250.
* `V`: The number of dimensions (i.e. vocabulary size) per sample. In the manuscript, this is set to 1000.
* `id`: A descriptive ID to associate with any saved results.
* `method`: The alignment strategy to pass to `align_topics`.
* `n_models`: The total number of models to fit to the simulated data. In the
manuscript, this is set to 10.
* `out_dir`: If results are saved, where should they be saved to?
* `save`: Should any results be saved?

The packages used in this vignette are given below. Note that we also load
simulation functions from an external github repository (not this package). This
provides the `simulate_lda` function used to generate the data for this example.

```{r}
library(MCMCpack)
library(alto)
library(dplyr)
library(ggplot2)
library(purrr)
library(readr)
library(stringr)
source("https://raw.githubusercontent.com/krisrs1128/topic_align/main/simulations/simulation_functions.R")
my_theme()
```

The block below simulates data `x` according to an LDA models with parameters
specified above. The topics are relatively sparse, with $\lambda_{\beta} = 0.1$
and $\lambda_{\gamma} = 0.5$. Each sample has 10,000 counts.

```{r}
attach(params)
lambdas <- list(beta = 0.1, gamma = .5, count = 1e4)
betas <- rdirichlet(K, rep(lambdas$beta, V))
gammas <- rdirichlet(N, rep(lambdas$gamma, K))
x <- simulate_lda(betas, gammas, lambda = lambdas$count)
```

Next, we run the LDA models and compute the alignment.

```{r}
lda_params <- map(1:n_models, ~ list(k = .))
names(lda_params) <- str_c("K", 1:n_models)
alignment <- x %>%
  run_lda_models(lda_params, reset = TRUE) %>%
  align_topics(method = params$method)
```

We can plot the flow diagram and compare the height-weight words across topics.

```{r}
plot(alignment)
plot_beta(alignment, c(2, 4))
```

We can plot measures of topic quality across $m$. The introduction of a low
quality topic at $K = 4$ is consistent with the fact that there are only 3 true
topics in this data.

```{r}
ggplot(topics(alignment), aes(m, coherence)) +
  geom_point(alpha = 0.5)

ggplot(topics(alignment), aes(m, refinement)) +
  geom_point(alpha = 0.5)
```

We can also compute the number of key topics associated with the alignment.

```{r}
key_topics <- 
  compute_number_of_paths(alignment) %>%
  mutate(id = params$id)
```

Finally, we save all the results (if wanted). The distinct IDs allow us to
gather alignments from across many replicates, and these are what are shown in
teh simulation section of the manuscript.

```{r}
id_vars <- params[c("out_dir", "method", "id", "N", "V", "K")]
scores <- topics(alignment) %>%
  mutate(id = params$id)

if (params$save) {
  dir.create(params$out_dir, recursive = TRUE)
  write_csv(key_topics, save_str("key_topics", id_vars))
  write_csv(scores, save_str("topics", id_vars))
  exper <- list(x, betas, gammas, alignment)
  save(exper, file = save_str("exper", id_vars, "rda"))
}
```
