---
title: "Measuring Runtimes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Measuring Runtimes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  K: 5
  V: 50
  N: 30
  n_models: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```

This vignette measures the runtime of a few steps in the alignment workflow.
Running this vignette with $V = 1000$, $N = 250$ gives the
estimates reported in the accompanying manuscript.

```{r}
library(MCMCpack)
library(alto)
library(dplyr)
library(purrr)
library(stringr)
library(tictoc)
source("https://raw.githubusercontent.com/krisrs1128/topic_align/main/simulations/simulation_functions.R")
```

For this simulation, we work with simulated LDA data, as in the "Identifying
True Topics" vignette.

```{r}
attach(params)
lambdas <- list(beta = 0.1, gamma = .5, count = 1e4)
betas <- rdirichlet(K, rep(lambdas$beta, V))
gammas <- rdirichlet(N, rep(lambdas$gamma, K))
x <- simulate_lda(betas, gammas, lambda = lambdas$count)
```

We split model running and alignment, so we can measure the computation times
separately. We use the `tictoc` library for this. In general, running the LDA
models consumes the majority of the time in an alignment workflow, especially
when the sample or vocabulary size is large.

```{r}
lda_params <- map(1:n_models, ~ list(k = .))
names(lda_params) <- str_c("K", 1:n_models)

tic()
lda_models <- run_lda_models(x, lda_params, reset = TRUE)
toc()

tic()
align_topics(lda_models, method = "product")
toc()

tic()
align_topics(lda_models, method = "transport")
toc()
```
